---
title: "Convert_h5ad_Seurat_Object"
output: html_document
date: "2025-02-14"
---

```{r setup, include=FALSE}
library(Seurat)
library(schard)
data_dir <- "/Users/anishadatta/Documents/DAL_Lab/Manuscript_CMBE/Axl-Inhibition-TME-Composition/Data" # Location of data
```

schard places adata.X (log1p data) from the h5ad file into both the counts slot and the data slot of the Seurat object, so the raw counts matrix needs to be manually added to the Seurat object
```{r Load raw counts matrix and save as rds file include = FALSE}
raw_counts_matrix <- read.csv(file.path(data_dir,"01_counts_transposed.csv"))
raw_counts_matrix <- raw_counts_matrix[,-1] #remove extra column corresponding to row number
saveRDS(raw_counts_matrix, file.path(data_dir, "01_counts_transposed.rds"))
```

```{r Convert raw counts matrix to Seurat object and save it include = FALSE}
#raw_counts_matrix <- readRDS(file.path(data_dir, "01_counts_transposed.rds"))
raw_counts <- CreateSeuratObject(counts = raw_counts_matrix)
saveRDS(raw_counts, file.path(data_dir, "01_raw_counts_seuratobj.rds"))
```

```{r Load and convert integrated h5ad file to Seurat object and save it include = FALSE}
h5ad_file <- paste0(data_dir, "/20250214-ADatta-Harmony-integrated-HVGonly-batch2only.h5ad")
integrated <- h5ad2seurat(h5ad_file)
saveRDS(integrated, file.path(data_dir, "01_integrated_seuratobj_needs_raw_counts.rds"))
```

```{r Replace log1p "counts" with actual raw counts matrix and save the corrected Seurat object include = FALSE}
#raw_counts <- readRDS(file.path(data_dir, "01_raw_counts_seuratobj.rds"))
#integrated <- readRDS(file.path(data_dir, "01_integrated_seuratobj_needs_raw_counts.rds"))
integrated@assays[["RNA"]]@counts@x <- raw_counts@assays[["RNA"]]@layers[["counts"]]@x
saveRDS(integrated, file.path(data_dir, "01_integrated_seuratobj.rds"))
```

> sessionInfo()
R version 4.4.1 (2024-06-14)
Platform: x86_64-apple-darwin20
Running under: macOS 15.7.7

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-x86_64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] schard_0.0.1       Seurat_5.1.0       SeuratObject_5.0.2 sp_2.1-4          

loaded via a namespace (and not attached):
  [1] spatstat.sparse_3.1-0       fs_1.6.5                    matrixStats_1.4.1          
  [4] lubridate_1.9.3             httr_1.4.7                  RColorBrewer_1.1-3         
  [7] doParallel_1.0.17           tools_4.4.1                 sctransform_0.4.1          
 [10] backports_1.5.0             utf8_1.2.4                  R6_2.5.1                   
 [13] uwot_0.2.2                  lazyeval_0.2.2              GetoptLong_1.0.5           
 [16] withr_3.0.2                 prettyunits_1.2.0           gridExtra_2.3              
 [19] progressr_0.15.0            cli_3.6.3                   Biobase_2.64.0             
 [22] spatstat.explore_3.3-3      fastDummies_1.7.4           spatstat.data_3.1-4        
 [25] S7_0.2.0                    readr_2.1.5                 ggridges_0.5.6             
 [28] pbapply_1.7-2               R.utils_2.13.0              parallelly_1.39.0          
 [31] sessioninfo_1.2.2           limma_3.60.6                readxl_1.4.3               
 [34] rstudioapi_0.17.1           RSQLite_2.3.8               generics_0.1.3             
 [37] shape_1.4.6.1               spatstat.random_3.3-2       ica_1.0-3                  
 [40] dplyr_1.1.4                 zip_2.3.1                   Matrix_1.7-1               
 [43] fansi_1.0.6                 S4Vectors_0.42.1            logger_0.4.0               
 [46] abind_1.4-8                 R.methodsS3_1.8.2           lifecycle_1.0.4            
 [49] yaml_2.3.10                 edgeR_4.2.2                 SummarizedExperiment_1.34.0
 [52] SparseArray_1.4.8           Rtsne_0.17                  grid_4.4.1                 
 [55] blob_1.2.4                  promises_1.3.0              dqrng_0.4.1                
 [58] crayon_1.5.3                dir.expiry_1.10.0           miniUI_0.1.1.1             
 [61] lattice_0.22-6              beachmat_2.20.0             cowplot_1.1.3              
 [64] chromote_0.5.1              pillar_1.9.0                knitr_1.49                 
 [67] ComplexHeatmap_2.18.0       metapod_1.10.1              GenomicRanges_1.56.2       
 [70] rjson_0.2.23                future.apply_1.11.3         codetools_0.2-20           
 [73] leiden_0.4.3.1              glue_1.8.0                  spatstat.univar_3.1-1      
 [76] data.table_1.16.2           vctrs_0.6.5                 png_0.1-8                  
 [79] spam_2.11-0                 cellranger_1.1.0            gtable_0.3.6               
 [82] cachem_1.1.0                OmnipathR_3.15.9            xfun_0.49                  
 [85] S4Arrays_1.4.1              mime_0.12                   liana_0.1.14               
 [88] survival_3.7-0              SingleCellExperiment_1.26.0 iterators_1.0.14           
 [91] statmod_1.5.0               bluster_1.12.0              fitdistrplus_1.2-1         
 [94] ROCR_1.0-11                 nlme_3.1-166                bit64_4.5.2                
 [97] progress_1.2.3              filelock_1.0.3              RcppAnnoy_0.0.22           
[100] GenomeInfoDb_1.40.1         irlba_2.3.5.1               KernSmooth_2.23-24         
[103] colorspace_2.1-1            BiocGenerics_0.50.0         DBI_1.2.3                  
[106] tidyselect_1.2.1            processx_3.8.6              bit_4.5.0                  
[109] compiler_4.4.1              curl_6.0.1                  rvest_1.0.4                
[112] httr2_1.0.6                 BiocNeighbors_1.22.0        basilisk.utils_1.19.1      
[115] xml2_1.3.6                  DelayedArray_0.30.1         plotly_4.10.4              
[118] checkmate_2.3.2             scales_1.4.0                lmtest_0.9-40              
[121] rappdirs_0.3.3              goftest_1.2-3               stringr_1.5.1              
[124] digest_0.6.37               spatstat.utils_3.1-1        rmarkdown_2.29             
[127] basilisk_1.19.1             XVector_0.44.0              htmltools_0.5.8.1          
[130] pkgconfig_2.0.3             sparseMatrixStats_1.16.0    MatrixGenerics_1.16.0      
[133] fastmap_1.2.0               rlang_1.1.4                 GlobalOptions_0.1.2        
[136] htmlwidgets_1.6.4           UCSC.utils_1.0.0            shiny_1.9.1                
[139] DelayedMatrixStats_1.26.0   farver_2.1.2                zoo_1.8-12                 
[142] jsonlite_1.8.9              BiocParallel_1.38.0         R.oo_1.27.0                
[145] BiocSingular_1.20.0         magrittr_2.0.3              scuttle_1.14.0             
[148] GenomeInfoDbData_1.2.12     dotCall64_1.2               patchwork_1.3.0            
[151] Rcpp_1.0.13-1               reticulate_1.40.0           stringi_1.8.4              
[154] zlibbioc_1.50.0             MASS_7.3-61                 plyr_1.8.9                 
[157] parallel_4.4.1              listenv_0.9.1               ggrepel_0.9.6              
[160] deldir_2.0-4                splines_4.4.1               tensor_1.5                 
[163] hms_1.1.3                   circlize_0.4.16             locfit_1.5-9.10            
[166] ps_1.8.1                    igraph_2.1.1                spatstat.geom_3.3-4        
[169] RcppHNSW_0.6.0              reshape2_1.4.4              stats4_4.4.1               
[172] ScaledMatrix_1.12.0         XML_3.99-0.17               evaluate_1.0.1             
[175] scran_1.32.0                tzdb_0.4.0                  foreach_1.5.2              
[178] httpuv_1.6.15               polyclip_1.10-7             RANN_2.6.2                 
[181] tidyr_1.3.1                 purrr_1.0.2                 future_1.34.0              
[184] clue_0.3-66                 scattermore_1.2             ggplot2_4.0.0              
[187] rsvd_1.0.5                  xtable_1.8-6                RSpectra_0.16-2            
[190] later_1.3.2                 viridisLite_0.4.2           tibble_3.2.1               
[193] websocket_1.4.4             memoise_2.0.1               IRanges_2.38.1             
[196] cluster_2.1.6               timechange_0.3.0            globals_0.16.3   