---
title: "ClusteringandAnnotation"
output: html_document
date: "2025-02-14"
---

```{r setup, include=FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(clustree)
library(writexl)
library(readxl)
library(stringr)
data_dir <- "/Users/anishadatta/Documents/DAL_Lab/Manuscript_CMBE/Axl-Inhibition-TME-Composition/Data" # Location of data
```

```{r Load intgerated Seurat object with correct raw counts matrix include = FALSE}
integrated_data <- readRDS(file.path(data_dir, "01_integrated_seuratobj.rds"))
```

```{r Create NN graphs for the integrated Seurat object and determine the optimal clustering resolution parameter include = FALSE}
# Find nearest neighbors and generate NN/SNN graphs (on Harmony-integrated PC reduction)
integrated_data <- FindNeighbors(integrated_data, reduction = "Xpcaharmony_") 

# Find clusters on SNN graph
res = c(0.01, 0.05, 0.1, 0.15, 0.16, 0.17, 0.18, 0.19, 0.2, 0.21, 0.22, 0.23, 0.24, 0.25, 0.3, 0.4, 0.5, 0.75, 1)
count_matrix <- GetAssayData(integrated_data, layer = "counts")
num_cells <- ncol(count_matrix)
res_clust_id_df <- data.frame(col1 = numeric(num_cells)) # create a dataframe to store each cell's cluster identity for each resolution parameter
for (x in res){
  cluster <- FindClusters(integrated_data, resolution = x)
  new_res <- paste0("res_",x)
  new_res_data <- cluster@meta.data$seurat_clusters
  res_clust_id_df[[new_res]] <- new_res_data
}

for (x in res){
  new_res <- paste0("res_",x)
  cluster@meta.data[[new_res]] <- res_clust_id_df[new_res] 
}
# https://cran.r-project.org/web/packages/clustree/vignettes/clustree.html
pdf(file.path(data_dir,"02_ClusterTree.pdf"), width = 10, height = 20)
clustree(cluster, prefix = "res_")
dev.off()
```

```{r Create clustered Seurat object with optimal resolution parameter and identify DEGs for the clusters to aid cluster annotation include = FALSE}
clustered_data <- FindClusters(integrated_data, resolution = 0.22)

all_markers <- FindAllMarkers(clustered_data, only.pos = TRUE)
all_markers_filtered <- subset(all_markers, p_val_adj < 0.05) 
all_markers_filtered %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10) %>%
    ungroup() -> top10
top10_uniquegenes <- unique(top10$gene)
pdf(file.path(data_dir, "02_DotPlot_Cluster_DEGs.pdf"), width = 25, height = 10)
DotPlot(clustered_data, features = top10_uniquegenes, dot.scale = 8) + RotatedAxis()
dev.off()

cluster_upreg_genes <- all_markers_filtered %>%
  group_by(cluster) %>%
  dplyr::filter(avg_log2FC > 1) 
cluster_upreg_genes <- split(cluster_upreg_genes, cluster_upreg_genes$cluster)
write_xlsx(cluster_upreg_genes, file.path(data_dir, "02_upreg_markers_by_cluster.xlsx"))
```

```{r Annotate and assign cluster labels include = FALSE}
cluster_ids <- c("Macrophage 1", "Macrophage 2", "Macrophage 3", "Dendritic Cell", "Untreated Tumor Cell", "Injured Tumor Cell", "Incomplete Efferocytosis", "LAMP3 Immune Cells")
names(cluster_ids) <- levels(clustered_data)
clustered_data <- RenameIdents(clustered_data, cluster_ids)
annotations <- clustered_data@active.ident
clustered_data$seurat_annotations <- annotations 
saveRDS(clustered_data, file.path(data_dir, "02_annotated_clustered_seuratobj.rds"))
pdf(file.path(data_dir, "02_AnnotatedClusters.pdf"), width = 7, height = 8)
DimPlot(clustered_data, reduction = "Xumap_", label = FALSE) + coord_fixed(ratio = 1.25) 
dev.off()
```

```{r Generate dotplot with annotated cluster names include = FALSE}
pdf(file.path(data_dir, "02_DotPlot_Annotated_Cluster_DEGs.pdf"), width = 10, height = 25)
DotPlot(clustered_data, features = top10_uniquegenes, dot.scale = 8) + RotatedAxis() + coord_flip()
dev.off()
```

```{r Generate UMAPs labeled by experimental condition include = FALSE}
pdf(file.path(data_dir, "02_A375State.pdf"), width = 4.54, height = 5.29)
DimPlot(clustered_data, reduction = "Xumap_", group.by = "injured", label = FALSE, cols = c("#148740", "#FC0098")) + coord_fixed(ratio = 1.25)
dev.off()

pdf(file.path(data_dir, "02_Inhibitor.pdf"), width = 4.54, height = 5.29)
DimPlot(clustered_data, reduction = "Xumap_", group.by = "treat", label = FALSE, cols = c("#F8766D", "#00BFC4")) + coord_fixed(ratio = 1.25)
dev.off()

pdf(file.path(data_dir, "02_Culture.pdf"), width = 4.54, height = 5.29)
DimPlot(clustered_data, reduction = "Xumap_", group.by = "culture", label = FALSE, cols = c("#A510E0", "#41464F")) + coord_fixed(ratio = 1.25)
dev.off()
```

```{r Generate UMAP labeled by cell type include = FALSE}
clustered_data$three_cell_type <- ifelse(clustered_data$seurat_annotations %in% c("Macrophage 1", "Macrophage 2", "Macrophage 3"), "Mac", ifelse(clustered_data$seurat_annotations %in% c("Untreated Tumor Cell", "Injured Tumor Cell"), "TC", ifelse(clustered_data$seurat_annotations %in% c("Dendritic Cell"), "DC", "None")))

pdf(file.path(data_dir, "02_ThreeCellType.pdf"), width = 4.54 , height = 5.29)
DimPlot(clustered_data, reduction = "Xumap_", group.by = "three_cell_type", label = FALSE, cols = c("#ABA6E0", "#719CB9", "grey", "#5A5A5A")) + coord_fixed(ratio = 1.25)
dev.off()
```

```{r Generate and plot module scores for M1 and M2 macrophage gene signatures include = FALSE}
# Use gene signatures from Oshi et al. Scientific Reports. (2020) https://www.nature.com/articles/s41598-020-73624-w#Sec2
Oshi_Cell_Types <- read_excel(file.path(data_dir, "02_13059_2017_1349_MOESM3_ESM.xlsx"))

M1_gene_lists <- filter(Oshi_Cell_Types, str_detect(Celltype_Source_ID, "Macrophages M1"))
M1_genes <- M1_gene_lists[, -(1:2)]
M1_genes <- unlist(as.list(M1_genes))
M1_genes <- unique(M1_genes) # 188 genes + 1 NA
M1_genes <- M1_genes[-29] # remove the NA

M2_gene_lists <- filter(Oshi_Cell_Types, str_detect(Celltype_Source_ID, "Macrophages M2"))
M2_genes <- M2_gene_lists[, -(1:2)]
M2_genes <- unlist(as.list(M2_genes))
M2_genes <- unique(M2_genes) # 160 genes + 1 NA
M2_genes <- M2_genes[-41] # remove the NA

# Filter M1 genes and M2 genes to those actually in the Seurat object
HVGs <- rownames(clustered_data)
M1_genes_in_HVG <- intersect(M1_genes, HVGs) # 74 genes
M2_genes_in_HVG <- intersect(M2_genes, HVGs) # 50 genes 
# Calculate module scores
clustered_data <- AddModuleScore(object = clustered_data, features = list(c(M1_genes_in_HVG)), ctrl = 5, name = "M1 Signature")
clustered_data <- AddModuleScore(object = clustered_data, features = list(c(M2_genes_in_HVG)), ctrl = 5, name = "M2 Signature")
# Visualize module scores
pdf(file.path(data_dir, "02_M1_Signature.pdf"), width = 4.54 , height = 5.29)
FeaturePlot(clustered_data, features = "M1 Signature1", cols = c("blue", "red")) + coord_fixed(ratio = 1.25)
dev.off()
pdf(file.path(data_dir, "02_M2_Signature.pdf"), width = 4.54 , height = 5.29)
FeaturePlot(clustered_data, features = "M2 Signature1", cols = c("blue", "red")) + coord_fixed(ratio = 1.25)
dev.off()
```

> sessionInfo()
R version 4.4.1 (2024-06-14)
Platform: x86_64-apple-darwin20
Running under: macOS 15.7.7

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-x86_64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] stringr_1.5.1      readxl_1.4.3       writexl_1.5.1      clustree_0.5.1    
 [5] ggraph_2.2.1       ggplot2_4.0.0      patchwork_1.3.0    Seurat_5.1.0      
 [9] SeuratObject_5.0.2 sp_2.1-4           dplyr_1.1.4       

loaded via a namespace (and not attached):
  [1] deldir_2.0-4           pbapply_1.7-2          gridExtra_2.3         
  [4] rlang_1.1.4            magrittr_2.0.3         RcppAnnoy_0.0.22      
  [7] spatstat.geom_3.3-4    matrixStats_1.4.1      ggridges_0.5.6        
 [10] compiler_4.4.1         png_0.1-8              vctrs_0.6.5           
 [13] reshape2_1.4.4         pkgconfig_2.0.3        fastmap_1.2.0         
 [16] utf8_1.2.4             promises_1.3.0         purrr_1.0.2           
 [19] xfun_0.49              cachem_1.1.0           jsonlite_1.8.9        
 [22] goftest_1.2-3          later_1.3.2            tweenr_2.0.3          
 [25] spatstat.utils_3.1-1   irlba_2.3.5.1          parallel_4.4.1        
 [28] cluster_2.1.6          R6_2.5.1               ica_1.0-3             
 [31] stringi_1.8.4          RColorBrewer_1.1-3     spatstat.data_3.1-4   
 [34] reticulate_1.40.0      parallelly_1.39.0      spatstat.univar_3.1-1 
 [37] cellranger_1.1.0       lmtest_0.9-40          scattermore_1.2       
 [40] Rcpp_1.0.13-1          knitr_1.49             tensor_1.5            
 [43] future.apply_1.11.3    zoo_1.8-12             sctransform_0.4.1     
 [46] httpuv_1.6.15          Matrix_1.7-1           splines_4.4.1         
 [49] igraph_2.1.1           tidyselect_1.2.1       viridis_0.6.5         
 [52] rstudioapi_0.17.1      abind_1.4-8            spatstat.random_3.3-2 
 [55] codetools_0.2-20       miniUI_0.1.1.1         spatstat.explore_3.3-3
 [58] listenv_0.9.1          lattice_0.22-6         tibble_3.2.1          
 [61] plyr_1.8.9             withr_3.0.2            shiny_1.9.1           
 [64] S7_0.2.0               ROCR_1.0-11            evaluate_1.0.1        
 [67] Rtsne_0.17             future_1.34.0          fastDummies_1.7.4     
 [70] survival_3.7-0         polyclip_1.10-7        fitdistrplus_1.2-1    
 [73] pillar_1.9.0           KernSmooth_2.23-24     plotly_4.10.4         
 [76] generics_0.1.3         RcppHNSW_0.6.0         scales_1.4.0          
 [79] globals_0.16.3         xtable_1.8-6           glue_1.8.0            
 [82] lazyeval_0.2.2         tools_4.4.1            data.table_1.16.2     
 [85] RSpectra_0.16-2        RANN_2.6.2             graphlayouts_1.2.1    
 [88] leiden_0.4.3.1         dotCall64_1.2          tidygraph_1.3.1       
 [91] cowplot_1.1.3          grid_4.4.1             tidyr_1.3.1           
 [94] colorspace_2.1-1       nlme_3.1-166           ggforce_0.4.2         
 [97] cli_3.6.3              spatstat.sparse_3.1-0  spam_2.11-0           
[100] fansi_1.0.6            viridisLite_0.4.2      uwot_0.2.2            
[103] gtable_0.3.6           digest_0.6.37          progressr_0.15.0      
[106] ggrepel_0.9.6          htmlwidgets_1.6.4      farver_2.1.2          
[109] memoise_2.0.1          htmltools_0.5.8.1      lifecycle_1.0.4       
[112] httr_1.4.7             mime_0.12              MASS_7.3-61