---
title: "PseudobulkandDESeq2"
output: html_document
date: "2025-02-14"
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(writexl)
library(DESeq2)
data_dir <- "/Users/anishadatta/Documents/DAL_Lab/Manuscript_CMBE/Axl-Inhibition-TME-Composition/Data" # Location of data
```

```{r Load annotated and clustered Seurat object include = FALSE}
annotated_clusters <- readRDS(file.path(data_dir, "02_annotated_clustered_seuratobj.rds"))
```

```{r Create Seurat object with active.ident reflecting culture, injury, treatment, and cluster include = FALSE}
annotated_clusters_stim <- annotated_clusters
annotated_clusters_stim$cluster_stim <- paste(annotated_clusters_stim$culture, annotated_clusters_stim$injured, annotated_clusters_stim$treat, annotated_clusters_stim$seurat_annotations, sep ="_")
Idents(annotated_clusters_stim) <- "cluster_stim"
```

```{r Pseudobulk prior to DESeq2 include = FALSE}
# First determine which clusters can be pseudobulked for each donor: min cells = 30
cell_counts_per_donor <- table(annotated_clusters_stim@meta.data[["cluster_stim"]], annotated_clusters_stim@meta.data[["donor"]])

write_xlsx(as.data.frame(cell_counts_per_donor), file.path(data_dir, "05_num_single_cells_per_donor_per_cluster.xlsx"))

meets_threshold <- apply(cell_counts_per_donor, 2, function(x) names(x[x >= 30]))

cells_to_include <- NULL
for (donor in names(meets_threshold)) {
  for (cluster in meets_threshold[[donor]]) {
    # Get the cells for this cluster and donor
    cells_to_include <- c(cells_to_include, colnames(annotated_clusters_stim)[annotated_clusters_stim@meta.data$cluster_stim == cluster & annotated_clusters_stim@meta.data$donor == donor])
  }
}

cells_to_pseudobulk <- subset(annotated_clusters_stim, cells = cells_to_include)

pseudo_bulk <- AggregateExpression(cells_to_pseudobulk, assays = "RNA", return.seurat = TRUE, group.by = c("culture", "injured", "treat", "donor", "seurat_annotations"))

# Remove donor from active ident so that there is more than 1 "cell" per ident 
pseudo_bulk$cluster_stim <- paste(pseudo_bulk$culture, pseudo_bulk$injured, pseudo_bulk$treat, pseudo_bulk$seurat_annotations, sep = "_")
Idents(pseudo_bulk) <- "cluster_stim"

saveRDS(pseudo_bulk, file.path(data_dir, "05_pseudobulked_data_seuratobj.rds"))

donor_counts <- as.data.frame(table(pseudo_bulk@active.ident))
write_xlsx(donor_counts, file.path(data_dir, "05_num_donors_per_pseudobulked_cluster.xlsx"))
```

```{r DESeq2 on inhibitor treatment: Bem vs DMSO vehicle include = FALSE}
# List of comparisons to be made
ident1 <- c("co_N_Bem_Macrophage 3", "co_N_Bem_Untreated Tumor Cell", "co_Y_Bem_Macrophage 1", "co_Y_Bem_Injured Tumor Cell", "tri_N_Bem_Dendritic Cell", "tri_N_Bem_Macrophage 2", "tri_N_Bem_Untreated Tumor Cell", "tri_Y_Bem_Dendritic Cell", "tri_Y_Bem_Macrophage 2",  "tri_Y_Bem_Injured Tumor Cell")
ident2 <- c("co_N_DMSO_Macrophage 3", "co_N_DMSO_Untreated Tumor Cell", "co_Y_DMSO_Macrophage 1", "co_Y_DMSO_Injured Tumor Cell", "tri_N_DMSO_Dendritic Cell", "tri_N_DMSO_Macrophage 2", "tri_N_DMSO_Untreated Tumor Cell", "tri_Y_DMSO_Dendritic Cell", "tri_Y_DMSO_Macrophage 2",  "tri_Y_DMSO_Injured Tumor Cell")
comparisons <- list(ident1 = ident1, ident2 = ident2)
num_comparisons <- length(comparisons$ident1)

# Create dataframe to store DESeq2 results for each comparison
not_adjusted_results <- data.frame(baseMean = numeric(0), 
                                   log2FoldChange = numeric(0), 
                                   lfcSE = numeric(0), 
                                   stat = numeric(0), 
                                   pvalue = numeric(0), 
                                   padj = numeric(0),
                                   ident1 = character(0), 
                                   ident2 = character(0), 
                                   gene = character(0), 
                                   ComparisonNumber = numeric(0),
                                   stringsAsFactors = FALSE)
# Loop through the comparisons of interest and run DESeq2
for (i in 1:num_comparisons){
  id1 = comparisons$ident1[i]
  id2 = comparisons$ident2[i]
  comparison <- subset(pseudo_bulk, idents = c(id1, id2))
  comparison_count <- GetAssayData(comparison, layer = "counts")
  comparison_meta_data <- comparison@meta.data[, c("donor","treat")]
  comparison_meta_data$donor <- factor(comparison_meta_data$donor)
  comparison_meta_data$treat <- factor(comparison_meta_data$treat)

  dds_treat <- DESeqDataSetFromMatrix(countData = comparison_count, colData = comparison_meta_data, design = ~treat + donor)
  dds_treat <- DESeq(dds_treat)
  res_treat <- results(dds_treat, contrast = c("treat", "Bem", "DMSO"))
  res_treat_df <- as.data.frame(res_treat)
  
  res_treat_df$ident1 <- id1
  res_treat_df$ident2 <- id2
  res_treat_df$gene <- rownames(res_treat_df)
  res_treat_df$ComparisonNumber <- i
  not_adjusted_results <- rbind(not_adjusted_results, res_treat_df)
}

# BH correct across all the pairwise comparisons made in the question of Bem vs DMSO vehicle
adjusted_results <- not_adjusted_results
adjusted_results$padj <- p.adjust(adjusted_results$pvalue, method = "BH")
saveRDS(adjusted_results, file.path(data_dir, "05_DESeq2_results_inhibitor_treatment_df.rds"))
# Also save as an Excel file such that each sheet contains the results from one pairwise comparison
split_adjusted_results_by_comparison <- split(adjusted_results, adjusted_results$ComparisonNumber)
write_xlsx(split_adjusted_results_by_comparison, file.path(data_dir, "05_DESeq2_results_inhibitor_treatment.xlsx"))
```

```{r DESeq2 on A375 state: Injured vs Untreated include = FALSE}
# List of comparisons to be made
ident1 <- c("co_Y_DMSO_Macrophage 1", "co_Y_DMSO_Injured Tumor Cell", "co_Y_Bem_Macrophage 1", "co_Y_Bem_Injured Tumor Cell", "tri_Y_DMSO_Macrophage 2", "tri_Y_DMSO_Dendritic Cell", "tri_Y_DMSO_Injured Tumor Cell", "tri_Y_Bem_Macrophage 2", "tri_Y_Bem_Dendritic Cell", "tri_Y_Bem_Injured Tumor Cell")
ident2 <- c("co_N_DMSO_Macrophage 3", "co_N_DMSO_Untreated Tumor Cell", "co_N_Bem_Macrophage 3", "co_N_Bem_Untreated Tumor Cell", "tri_N_DMSO_Macrophage 2", "tri_N_DMSO_Dendritic Cell", "tri_N_DMSO_Untreated Tumor Cell", "tri_N_Bem_Macrophage 2", "tri_N_Bem_Dendritic Cell", "tri_N_Bem_Untreated Tumor Cell")
comparisons <- list(ident1 = ident1, ident2 = ident2)
num_comparisons <- length(comparisons$ident1)

# Create dataframe to store DESeq2 results for each comparison
not_adjusted_results <- data.frame(baseMean = numeric(0), 
                                   log2FoldChange = numeric(0), 
                                   lfcSE = numeric(0), 
                                   stat = numeric(0), 
                                   pvalue = numeric(0), 
                                   padj = numeric(0),
                                   ident1 = character(0), 
                                   ident2 = character(0), 
                                   gene = character(0), 
                                   ComparisonNumber = numeric(0),
                                   stringsAsFactors = FALSE)
# Loop through the comparisons of interest and run DESeq2
for (i in 1:num_comparisons){
  id1 = comparisons$ident1[i]
  id2 = comparisons$ident2[i]
  comparison <- subset(pseudo_bulk, idents = c(id1, id2))
  comparison_count <- GetAssayData(comparison, layer = "counts")
  comparison_meta_data <- comparison@meta.data[, c("donor","injured")]
  comparison_meta_data$donor <- factor(comparison_meta_data$donor)
  comparison_meta_data$injured <- factor(comparison_meta_data$injured)

  dds_injured <- DESeqDataSetFromMatrix(countData = comparison_count, colData = comparison_meta_data, design = ~injured + donor)
  dds_injured <- DESeq(dds_injured)
  res_injured <- results(dds_injured, contrast = c("injured", "Y", "N"))
  res_injured_df <- as.data.frame(res_injured)
  
  res_injured_df$ident1 <- id1
  res_injured_df$ident2 <- id2
  res_injured_df$gene <- rownames(res_injured_df)
  res_injured_df$ComparisonNumber <- i
  not_adjusted_results <- rbind(not_adjusted_results, res_injured_df)
}
# BH correct across all the pairwise comparisons made in the question of Injured vs Untreated
adjusted_results <- not_adjusted_results
adjusted_results$padj <- p.adjust(adjusted_results$pvalue, method = "BH")
saveRDS(adjusted_results, file.path(data_dir, "05_DESeq2_results_A375_state_df.rds"))
# Also save as an Excel file such that each sheet contains the results from one pairwise comparison
split_adjusted_results_by_comparison <- split(adjusted_results, adjusted_results$ComparisonNumber)
write_xlsx(split_adjusted_results_by_comparison, file.path(data_dir, "05_DESeq2_results_A375_state.xlsx"))
```

```{r DESeq2 on culture: Tri-culture vs Co-culture include = FALSE}
# List of comparisons to be made
ident1 <- c("tri_N_DMSO_Macrophage 2", "tri_N_DMSO_Untreated Tumor Cell", "tri_N_Bem_Macrophage 2", "tri_N_Bem_Untreated Tumor Cell", "tri_Y_DMSO_Macrophage 2", "tri_Y_DMSO_Injured Tumor Cell", "tri_Y_Bem_Macrophage 2", "tri_Y_Bem_Injured Tumor Cell")
ident2 <- c("co_N_DMSO_Macrophage 3", "co_N_DMSO_Untreated Tumor Cell", "co_N_Bem_Macrophage 3", "co_N_Bem_Untreated Tumor Cell", "co_Y_DMSO_Macrophage 1", "co_Y_DMSO_Injured Tumor Cell", "co_Y_Bem_Macrophage 1", "co_Y_Bem_Injured Tumor Cell")
comparisons <- list(ident1 = ident1, ident2 = ident2)
num_comparisons <- length(comparisons$ident1)

# Create dataframe to store DESeq2 results for each comparison
not_adjusted_results <- data.frame(baseMean = numeric(0), 
                                   log2FoldChange = numeric(0), 
                                   lfcSE = numeric(0), 
                                   stat = numeric(0), 
                                   pvalue = numeric(0), 
                                   padj = numeric(0),
                                   ident1 = character(0), 
                                   ident2 = character(0), 
                                   gene = character(0), 
                                   ComparisonNumber = numeric(0),
                                   stringsAsFactors = FALSE)
# Loop through the comparisons of interest and run DESeq2
for (i in 1:num_comparisons){
  id1 = comparisons$ident1[i]
  id2 = comparisons$ident2[i]
  comparison <- subset(pseudo_bulk, idents = c(id1, id2))
  comparison_count <- GetAssayData(comparison, layer = "counts")
  comparison_meta_data <- comparison@meta.data[, c("donor","culture")]
  comparison_meta_data$donor <- factor(comparison_meta_data$donor)
  comparison_meta_data$culture <- factor(comparison_meta_data$culture)

  dds_culture <- DESeqDataSetFromMatrix(countData = comparison_count, colData = comparison_meta_data, design = ~culture + donor)
  dds_culture <- DESeq(dds_culture)
  res_culture <- results(dds_culture, contrast = c("culture", "tri", "co"))
  res_culture_df <- as.data.frame(res_culture)
  
  res_culture_df$ident1 <- id1
  res_culture_df$ident2 <- id2
  res_culture_df$gene <- rownames(res_culture_df)
  res_culture_df$ComparisonNumber <- i
  not_adjusted_results <- rbind(not_adjusted_results, res_culture_df)
}
# BH correct across all the pairwise comparisons made in the question of Tri-Culture vs Co-Culture
adjusted_results <- not_adjusted_results
adjusted_results$padj <- p.adjust(adjusted_results$pvalue, method = "BH")
saveRDS(adjusted_results, file.path(data_dir, "05_DESeq2_results_culture_type_df.rds"))
# Also save as an Excel file such that each sheet contains the results from one pairwise comparison
split_adjusted_results_by_comparison <- split(adjusted_results, adjusted_results$ComparisonNumber)
write_xlsx(split_adjusted_results_by_comparison, file.path(data_dir, "05_DESeq2_results_culture_type.xlsx"))
```

> sessionInfo()
R version 4.4.1 (2024-06-14)
Platform: x86_64-apple-darwin20
Running under: macOS 15.7.7

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-x86_64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] DESeq2_1.44.0               SummarizedExperiment_1.34.0 Biobase_2.64.0             
 [4] MatrixGenerics_1.16.0       matrixStats_1.4.1           GenomicRanges_1.56.2       
 [7] GenomeInfoDb_1.40.1         IRanges_2.38.1              S4Vectors_0.42.1           
[10] BiocGenerics_0.50.0         writexl_1.5.1               ggplot2_4.0.0              
[13] patchwork_1.3.0             Seurat_5.1.0                SeuratObject_5.0.2         
[16] sp_2.1-4                    tidyr_1.3.1                 dplyr_1.1.4                

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3      rstudioapi_0.17.1       jsonlite_1.8.9         
  [4] magrittr_2.0.3          spatstat.utils_3.1-1    farver_2.1.2           
  [7] zlibbioc_1.50.0         vctrs_0.6.5             ROCR_1.0-11            
 [10] spatstat.explore_3.3-3  S4Arrays_1.4.1          htmltools_0.5.8.1      
 [13] SparseArray_1.4.8       sctransform_0.4.1       parallelly_1.39.0      
 [16] KernSmooth_2.23-24      htmlwidgets_1.6.4       ica_1.0-3              
 [19] plyr_1.8.9              plotly_4.10.4           zoo_1.8-12             
 [22] igraph_2.1.1            mime_0.12               lifecycle_1.0.4        
 [25] pkgconfig_2.0.3         Matrix_1.7-1            R6_2.5.1               
 [28] fastmap_1.2.0           GenomeInfoDbData_1.2.12 fitdistrplus_1.2-1     
 [31] future_1.34.0           shiny_1.9.1             digest_0.6.37          
 [34] colorspace_2.1-1        tensor_1.5              RSpectra_0.16-2        
 [37] irlba_2.3.5.1           progressr_0.15.0        fansi_1.0.6            
 [40] spatstat.sparse_3.1-0   httr_1.4.7              polyclip_1.10-7        
 [43] abind_1.4-8             compiler_4.4.1          withr_3.0.2            
 [46] S7_0.2.0                BiocParallel_1.38.0     fastDummies_1.7.4      
 [49] MASS_7.3-61             DelayedArray_0.30.1     tools_4.4.1            
 [52] lmtest_0.9-40           httpuv_1.6.15           future.apply_1.11.3    
 [55] goftest_1.2-3           glue_1.8.0              nlme_3.1-166           
 [58] promises_1.3.0          grid_4.4.1              Rtsne_0.17             
 [61] cluster_2.1.6           reshape2_1.4.4          generics_0.1.3         
 [64] gtable_0.3.6            spatstat.data_3.1-4     data.table_1.16.2      
 [67] utf8_1.2.4              XVector_0.44.0          spatstat.geom_3.3-4    
 [70] RcppAnnoy_0.0.22        ggrepel_0.9.6           RANN_2.6.2             
 [73] pillar_1.9.0            stringr_1.5.1           spam_2.11-0            
 [76] RcppHNSW_0.6.0          later_1.3.2             splines_4.4.1          
 [79] lattice_0.22-6          survival_3.7-0          deldir_2.0-4           
 [82] tidyselect_1.2.1        locfit_1.5-9.10         miniUI_0.1.1.1         
 [85] pbapply_1.7-2           knitr_1.49              gridExtra_2.3          
 [88] scattermore_1.2         xfun_0.49               stringi_1.8.4          
 [91] UCSC.utils_1.0.0        lazyeval_0.2.2          evaluate_1.0.1         
 [94] codetools_0.2-20        tibble_3.2.1            cli_3.6.3              
 [97] uwot_0.2.2              xtable_1.8-6            reticulate_1.40.0      
[100] Rcpp_1.0.13-1           globals_0.16.3          spatstat.random_3.3-2  
[103] png_0.1-8               spatstat.univar_3.1-1   parallel_4.4.1         
[106] dotCall64_1.2           listenv_0.9.1           viridisLite_0.4.2      
[109] scales_1.4.0            ggridges_0.5.6          crayon_1.5.3           
[112] leiden_0.4.3.1          purrr_1.0.2             rlang_1.1.4            
[115] cowplot_1.1.3 