---
title: "GSEA"
output: html_document
date: "2025-02-14"
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(Seurat)
library(writexl)
library(readxl)
library(fgsea)
library(msigdbr)

library(patchwork)
library(ggplot2)
data_dir <- "/Users/anishadatta/Documents/DAL_Lab/Manuscript_CMBE/Axl-Inhibition-TME-Composition/Data" # Location of data
```

```{r GSEA on pairwise comparisons of inhibitor treatment: Bem vs DMSO include = FALSE}
# List of comparisons
ident1 <- c("co_N_Bem_Macrophage 3", "co_N_Bem_Untreated Tumor Cell", "co_Y_Bem_Macrophage 1", "co_Y_Bem_Injured Tumor Cell", "tri_N_Bem_Dendritic Cell", "tri_N_Bem_Macrophage 2", "tri_N_Bem_Untreated Tumor Cell", "tri_Y_Bem_Dendritic Cell", "tri_Y_Bem_Macrophage 2",  "tri_Y_Bem_Injured Tumor Cell")
ident2 <- c("co_N_DMSO_Macrophage 3", "co_N_DMSO_Untreated Tumor Cell", "co_Y_DMSO_Macrophage 1", "co_Y_DMSO_Injured Tumor Cell", "tri_N_DMSO_Dendritic Cell", "tri_N_DMSO_Macrophage 2", "tri_N_DMSO_Untreated Tumor Cell", "tri_Y_DMSO_Dendritic Cell", "tri_Y_DMSO_Macrophage 2",  "tri_Y_DMSO_Injured Tumor Cell")
comparisons <- list(ident1 = ident1, ident2 = ident2)
num_comparisons <- length(comparisons$ident1)

# Load DESeq2 results 
DE_res <- readRDS(file.path(data_dir, "05_DESeq2_results_inhibitor_treatment_df.rds"))

# Load Hallmark gene sets 
msigdb_gene_sets <- msigdbr(species = "Homo sapiens", category = "H")
gene_sets <- split(msigdb_gene_sets$gene_symbol, msigdb_gene_sets$gs_name)
 
# Generate dataframe to store GSEA results
GSEA_res_df <- data.frame(GeneSet = character(0), 
                                   p_adj = numeric(0),
                                   NES = numeric(0),
                                   size = numeric(0),
                                   LeadingEdgeGenes = character(0), 
                                   ident1 = character(0), 
                                   ident2 = character(0),
                                   ComparisonNumber = numeric(0))

for (k in 1:num_comparisons){
  id1 = comparisons$ident1[k]
  id2 = comparisons$ident2[k]
  current_comp <- paste0(id1, " vs ", id2)
  fgsea_data <- subset(DE_res, ComparisonNumber == k)
  fgsea_data <- fgsea_data %>% drop_na()
  row.names(fgsea_data) <- fgsea_data$gene
  ranks <- fgsea_data$stat
  names(ranks) <- rownames(fgsea_data)
  set.seed(20)
  fgsea_results <- fgsea(pathways = gene_sets,
                       stats = ranks, 
                       eps = 0,
                       minSize = 15, # min size of pathway overlap with genes in the data
                       maxSize = 500)

  #https://stephenturner.github.io/deseq-to-fgsea/
  fgsea_results <- fgsea_results %>%
    as_tibble() %>%
    arrange(desc(NES))
  
  p_val <- fgsea_results$padj
  num_genes_in_set <- fgsea_results$size #size of gene set after removing genes not in the data
  nes <- fgsea_results$NES
  leading_edge <- fgsea_results$leadingEdge
  for (l in 1:length(fgsea_results$leadingEdge)){
    GeneSetName <- fgsea_results$pathway[l]
    LeadingEdgeGeneNames <- paste(leading_edge[[l]], collapse = ", ")
    GSEA_res_df <- rbind(GSEA_res_df, data.frame(GeneSet = GeneSetName, p_adj = p_val[[l]], NES = nes[[l]], size = num_genes_in_set[[l]], LeadingEdgeGenes = LeadingEdgeGeneNames, ident1 = id1, ident2 = id2, ComparisonNumber = k))
  }
}

# Save GSEA results dataframe 
saveRDS(GSEA_res_df, file.path(data_dir, "06_GSEA_full_results_inhibitor_treatment_df.rds"))

# Save GSEA results in excel file such that each sheet contains the results for each comparison
split_by_comparison <- split(GSEA_res_df, GSEA_res_df$ComparisonNumber)
write_xlsx(split_by_comparison, file.path(data_dir, "06_GSEA_full_results_inhibitor_treatment.xlsx"))

# Subset to only signficant results
GSEA_res_df_hits <- subset(GSEA_res_df, p_adj < 0.05)

saveRDS(GSEA_res_df_hits, file.path(data_dir, "06_GSEA_significant_results_inhibitor_treatment_df.rds"))

split_hits_by_comparison <- split(GSEA_res_df_hits, GSEA_res_df_hits$ComparisonNumber)
write_xlsx(split_hits_by_comparison, file.path(data_dir, "06_GSEA_significant_results_inhibitor_treatment.xlsx"))
```

```{r Generate GSEA enrichment plot for Fig. 1C include = FALSE}
# List of comparisons
ident1 <- "co_N_Bem_Macrophage 3"
ident2 <- "co_N_DMSO_Macrophage 3"
comparisons <- list(ident1 = ident1, ident2 = ident2)
num_comparisons <- length(comparisons$ident1)

# Load DESeq2 results 
DE_res <- readRDS(file.path(data_dir, "05_DESeq2_results_inhibitor_treatment_df.rds"))

# Load Hallmark gene sets 
msigdb_gene_sets <- msigdbr(species = "Homo sapiens", category = "H")
gene_sets <- split(msigdb_gene_sets$gene_symbol, msigdb_gene_sets$gs_name)
 
# Generate dataframe to store GSEA results
GSEA_res_df <- data.frame(GeneSet = character(0), 
                                   p_adj = numeric(0),
                                   NES = numeric(0),
                                   size = numeric(0),
                                   LeadingEdgeGenes = character(0), 
                                   ident1 = character(0), 
                                   ident2 = character(0),
                                   ComparisonNumber = numeric(0))

for (k in 1:num_comparisons){
  id1 = comparisons$ident1[k]
  id2 = comparisons$ident2[k]
  current_comp <- paste0(id1, " vs ", id2)
  fgsea_data <- subset(DE_res, ComparisonNumber == k)
  fgsea_data <- fgsea_data %>% drop_na()
  row.names(fgsea_data) <- fgsea_data$gene
  ranks <- fgsea_data$stat
  names(ranks) <- rownames(fgsea_data)
  set.seed(20)
  fgsea_results <- fgsea(pathways = gene_sets,
                       stats = ranks, 
                       eps = 0,
                       minSize = 15, # min size of pathway overlap with genes in the data
                       maxSize = 500)

  #https://stephenturner.github.io/deseq-to-fgsea/
  fgsea_results <- fgsea_results %>%
    as_tibble() %>%
    arrange(desc(NES))
  
  p_val <- fgsea_results$padj
  num_genes_in_set <- fgsea_results$size #size of gene set after removing genes not in the data
  nes <- fgsea_results$NES
  leading_edge <- fgsea_results$leadingEdge
  for (l in 1:length(fgsea_results$leadingEdge)){
    GeneSetName <- fgsea_results$pathway[l]
    LeadingEdgeGeneNames <- paste(leading_edge[[l]], collapse = ", ")
    GSEA_res_df <- rbind(GSEA_res_df, data.frame(GeneSet = GeneSetName, p_adj = p_val[[l]], NES = nes[[l]], size = num_genes_in_set[[l]], LeadingEdgeGenes = LeadingEdgeGeneNames, ident1 = id1, ident2 = id2, ComparisonNumber = k))
  }
}

#https://bioconductor.org/packages/release/bioc/vignettes/fgsea/inst/doc/fgsea-tutorial.html
plotEnrichment(gene_sets[["HALLMARK_INTERFERON_ALPHA_RESPONSE"]],
               ranks) + labs(title="Interferon Alpha Response")
```

```{r GSEA on pairwise comparisons of A375 state: Injured vs Untreated include = FALSE}
# List of comparisons
ident1 <- c("co_Y_DMSO_Macrophage 1", "co_Y_DMSO_Injured Tumor Cell", "co_Y_Bem_Macrophage 1", "co_Y_Bem_Injured Tumor Cell", "tri_Y_DMSO_Macrophage 2", "tri_Y_DMSO_Dendritic Cell", "tri_Y_DMSO_Injured Tumor Cell", "tri_Y_Bem_Macrophage 2", "tri_Y_Bem_Dendritic Cell", "tri_Y_Bem_Injured Tumor Cell")
ident2 <- c("co_N_DMSO_Macrophage 3", "co_N_DMSO_Untreated Tumor Cell", "co_N_Bem_Macrophage 3", "co_N_Bem_Untreated Tumor Cell", "tri_N_DMSO_Macrophage 2", "tri_N_DMSO_Dendritic Cell", "tri_N_DMSO_Untreated Tumor Cell", "tri_N_Bem_Macrophage 2", "tri_N_Bem_Dendritic Cell", "tri_N_Bem_Untreated Tumor Cell")
comparisons <- list(ident1 = ident1, ident2 = ident2)
num_comparisons <- length(comparisons$ident1)

# Load DESeq2 results 
DE_res <- readRDS(file.path(data_dir, "05_DESeq2_results_A375_state_df.rds"))

# Load Hallmark gene sets 
msigdb_gene_sets <- msigdbr(species = "Homo sapiens", category = "H")
gene_sets <- split(msigdb_gene_sets$gene_symbol, msigdb_gene_sets$gs_name)
 
# Generate dataframe to store GSEA results
GSEA_res_df <- data.frame(GeneSet = character(0), 
                                   p_adj = numeric(0),
                                   NES = numeric(0),
                                   size = numeric(0),
                                   LeadingEdgeGenes = character(0), 
                                   ident1 = character(0), 
                                   ident2 = character(0),
                                   ComparisonNumber = numeric(0))

for (k in 1:num_comparisons){
  id1 = comparisons$ident1[k]
  id2 = comparisons$ident2[k]
  current_comp <- paste0(id1, " vs ", id2)
  fgsea_data <- subset(DE_res, ComparisonNumber == k)
  fgsea_data <- fgsea_data %>% drop_na()
  row.names(fgsea_data) <- fgsea_data$gene
  ranks <- fgsea_data$stat
  names(ranks) <- rownames(fgsea_data)
  set.seed(20)
  fgsea_results <- fgsea(pathways = gene_sets,
                       stats = ranks, 
                       eps = 0,
                       minSize = 15, # min size of pathway overlap with genes in the data
                       maxSize = 500)

  #https://stephenturner.github.io/deseq-to-fgsea/
  fgsea_results <- fgsea_results %>%
    as_tibble() %>%
    arrange(desc(NES))
  
  p_val <- fgsea_results$padj
  num_genes_in_set <- fgsea_results$size #size of gene set after removing genes not in the data
  nes <- fgsea_results$NES
  leading_edge <- fgsea_results$leadingEdge
  for (l in 1:length(fgsea_results$leadingEdge)){
    GeneSetName <- fgsea_results$pathway[l]
    LeadingEdgeGeneNames <- paste(leading_edge[[l]], collapse = ", ")
    GSEA_res_df <- rbind(GSEA_res_df, data.frame(GeneSet = GeneSetName, p_adj = p_val[[l]], NES = nes[[l]], size = num_genes_in_set[[l]], LeadingEdgeGenes = LeadingEdgeGeneNames, ident1 = id1, ident2 = id2, ComparisonNumber = k))
  }
}

# Save GSEA results dataframe 
saveRDS(GSEA_res_df, file.path(data_dir, "06_GSEA_full_results_A375_state_df.rds"))

# Save GSEA results in excel file such that each sheet contains the results for each comparison
split_by_comparison <- split(GSEA_res_df, GSEA_res_df$ComparisonNumber)
write_xlsx(split_by_comparison, file.path(data_dir, "06_GSEA_full_results_A375_state.xlsx"))

# Subset to only signficant results
GSEA_res_df_hits <- subset(GSEA_res_df, p_adj < 0.05)

saveRDS(GSEA_res_df_hits, file.path(data_dir, "06_GSEA_significant_results_A375_state_df.rds"))

split_hits_by_comparison <- split(GSEA_res_df_hits, GSEA_res_df_hits$ComparisonNumber)
write_xlsx(split_hits_by_comparison, file.path(data_dir, "06_GSEA_significant_results_A375_state.xlsx"))
```

```{r GSEA on pairwise comparisons of culture type: Tri-Culture vs Co-Culture include = FALSE}
# List of comparisons
ident1 <- c("tri_N_DMSO_Macrophage 2", "tri_N_DMSO_Untreated Tumor Cell", "tri_N_Bem_Macrophage 2", "tri_N_Bem_Untreated Tumor Cell", "tri_Y_DMSO_Macrophage 2", "tri_Y_DMSO_Injured Tumor Cell", "tri_Y_Bem_Macrophage 2", "tri_Y_Bem_Injured Tumor Cell")
ident2 <- c("co_N_DMSO_Macrophage 3", "co_N_DMSO_Untreated Tumor Cell", "co_N_Bem_Macrophage 3", "co_N_Bem_Untreated Tumor Cell", "co_Y_DMSO_Macrophage 1", "co_Y_DMSO_Injured Tumor Cell", "co_Y_Bem_Macrophage 1", "co_Y_Bem_Injured Tumor Cell")
comparisons <- list(ident1 = ident1, ident2 = ident2)
num_comparisons <- length(comparisons$ident1)

# Load DESeq2 results 
DE_res <- readRDS(file.path(data_dir, "05_DESeq2_results_culture_type_df.rds"))

# Load Hallmark gene sets 
msigdb_gene_sets <- msigdbr(species = "Homo sapiens", category = "H")
gene_sets <- split(msigdb_gene_sets$gene_symbol, msigdb_gene_sets$gs_name)
 
# Generate dataframe to store GSEA results
GSEA_res_df <- data.frame(GeneSet = character(0), 
                                   p_adj = numeric(0),
                                   NES = numeric(0),
                                   size = numeric(0),
                                   LeadingEdgeGenes = character(0), 
                                   ident1 = character(0), 
                                   ident2 = character(0),
                                   ComparisonNumber = numeric(0))

for (k in 1:num_comparisons){
  id1 = comparisons$ident1[k]
  id2 = comparisons$ident2[k]
  current_comp <- paste0(id1, " vs ", id2)
  fgsea_data <- subset(DE_res, ComparisonNumber == k)
  fgsea_data <- fgsea_data %>% drop_na()
  row.names(fgsea_data) <- fgsea_data$gene
  ranks <- fgsea_data$stat
  names(ranks) <- rownames(fgsea_data)
  set.seed(20)
  fgsea_results <- fgsea(pathways = gene_sets,
                       stats = ranks, 
                       eps = 0,
                       minSize = 15, # min size of pathway overlap with genes in the data
                       maxSize = 500)

  #https://stephenturner.github.io/deseq-to-fgsea/
  fgsea_results <- fgsea_results %>%
    as_tibble() %>%
    arrange(desc(NES))
  
  p_val <- fgsea_results$padj
  num_genes_in_set <- fgsea_results$size #size of gene set after removing genes not in the data
  nes <- fgsea_results$NES
  leading_edge <- fgsea_results$leadingEdge
  for (l in 1:length(fgsea_results$leadingEdge)){
    GeneSetName <- fgsea_results$pathway[l]
    LeadingEdgeGeneNames <- paste(leading_edge[[l]], collapse = ", ")
    GSEA_res_df <- rbind(GSEA_res_df, data.frame(GeneSet = GeneSetName, p_adj = p_val[[l]], NES = nes[[l]], size = num_genes_in_set[[l]], LeadingEdgeGenes = LeadingEdgeGeneNames, ident1 = id1, ident2 = id2, ComparisonNumber = k))
  }
}

# Save GSEA results dataframe 
saveRDS(GSEA_res_df, file.path(data_dir, "06_GSEA_full_results_culture_type_df.rds"))

# Save GSEA results in excel file such that each sheet contains the results for each comparison
split_by_comparison <- split(GSEA_res_df, GSEA_res_df$ComparisonNumber)
write_xlsx(split_by_comparison, file.path(data_dir, "06_GSEA_full_results_culture_type.xlsx"))

# Subset to only signficant results
GSEA_res_df_hits <- subset(GSEA_res_df, p_adj < 0.05)

saveRDS(GSEA_res_df_hits, file.path(data_dir, "06_GSEA_significant_results_culture_type_df.rds"))

split_hits_by_comparison <- split(GSEA_res_df_hits, GSEA_res_df_hits$ComparisonNumber)
write_xlsx(split_hits_by_comparison, file.path(data_dir, "06_GSEA_significant_results_culture_type.xlsx"))
```

> sessionInfo()
R version 4.4.1 (2024-06-14)
Platform: x86_64-apple-darwin20
Running under: macOS 15.7.7

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-x86_64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] ggplot2_4.0.0      patchwork_1.3.0    msigdbr_7.5.1      fgsea_1.30.0      
 [5] readxl_1.4.3       writexl_1.5.1      Seurat_5.1.0       SeuratObject_5.0.2
 [9] sp_2.1-4           tidyr_1.3.1        dplyr_1.1.4       

loaded via a namespace (and not attached):
  [1] deldir_2.0-4           pbapply_1.7-2          gridExtra_2.3         
  [4] rlang_1.1.4            magrittr_2.0.3         RcppAnnoy_0.0.22      
  [7] spatstat.geom_3.3-4    matrixStats_1.4.1      ggridges_0.5.6        
 [10] compiler_4.4.1         png_0.1-8              vctrs_0.6.5           
 [13] reshape2_1.4.4         stringr_1.5.1          pkgconfig_2.0.3       
 [16] fastmap_1.2.0          labeling_0.4.3         utf8_1.2.4            
 [19] promises_1.3.0         purrr_1.0.2            xfun_0.49             
 [22] jsonlite_1.8.9         goftest_1.2-3          later_1.3.2           
 [25] BiocParallel_1.38.0    spatstat.utils_3.1-1   irlba_2.3.5.1         
 [28] parallel_4.4.1         cluster_2.1.6          R6_2.5.1              
 [31] ica_1.0-3              stringi_1.8.4          RColorBrewer_1.1-3    
 [34] spatstat.data_3.1-4    reticulate_1.40.0      parallelly_1.39.0     
 [37] spatstat.univar_3.1-1  cellranger_1.1.0       lmtest_0.9-40         
 [40] scattermore_1.2        Rcpp_1.0.13-1          knitr_1.49            
 [43] tensor_1.5             future.apply_1.11.3    zoo_1.8-12            
 [46] sctransform_0.4.1      httpuv_1.6.15          Matrix_1.7-1          
 [49] splines_4.4.1          igraph_2.1.1           tidyselect_1.2.1      
 [52] abind_1.4-8            rstudioapi_0.17.1      spatstat.random_3.3-2 
 [55] codetools_0.2-20       miniUI_0.1.1.1         spatstat.explore_3.3-3
 [58] listenv_0.9.1          lattice_0.22-6         tibble_3.2.1          
 [61] plyr_1.8.9             withr_3.0.2            shiny_1.9.1           
 [64] S7_0.2.0               ROCR_1.0-11            evaluate_1.0.1        
 [67] Rtsne_0.17             future_1.34.0          fastDummies_1.7.4     
 [70] survival_3.7-0         polyclip_1.10-7        fitdistrplus_1.2-1    
 [73] pillar_1.9.0           KernSmooth_2.23-24     plotly_4.10.4         
 [76] generics_0.1.3         RcppHNSW_0.6.0         scales_1.4.0          
 [79] globals_0.16.3         xtable_1.8-6           glue_1.8.0            
 [82] lazyeval_0.2.2         tools_4.4.1            data.table_1.16.2     
 [85] RSpectra_0.16-2        babelgene_22.9         RANN_2.6.2            
 [88] leiden_0.4.3.1         dotCall64_1.2          fastmatch_1.1-4       
 [91] cowplot_1.1.3          grid_4.4.1             colorspace_2.1-1      
 [94] nlme_3.1-166           cli_3.6.3              spatstat.sparse_3.1-0 
 [97] spam_2.11-0            fansi_1.0.6            viridisLite_0.4.2     
[100] uwot_0.2.2             gtable_0.3.6           digest_0.6.37         
[103] progressr_0.15.0       ggrepel_0.9.6          htmlwidgets_1.6.4     
[106] farver_2.1.2           htmltools_0.5.8.1      lifecycle_1.0.4       
[109] httr_1.4.7             mime_0.12              MASS_7.3-61 